### 2013-05-22 doc/icml2013/algo.tex

%%% first moment
To warm up, let us consider linear regression
on the response $y$ given $x$.
From the model definition, we have $y = \beta_h^\top x + \epsilon$.
The challenge is that the regression coefficients $\beta_h$ depend on the random $h$.
%$\epsilon \sim \normal{0}{\E[\epsilon^2]}$, and $h$ is a random quantity
%independent of $x$.
The first key step is to average over this randomness by defining
average regression coefficients
$M_1 \eqdef \sum_{h=1}^k \pi_h \beta_h$.
Now we can express $y$ as a linear function of $x$ with non-random coefficients $M_1$
plus some a new noise term $\eta_1(x)$:
\begin{align}
  y &= \innerp{M_1}{x} +
  \underbrace{(\innerp{\beta_h - M_1}{x} + \epsilon)}_{\eqdef \eta_1(x)}. \label{eqn:y1}
\end{align}
The noise $\eta_1(x)$ is the sum of two terms:
(i) the \emph{mixing noise} $\innerp{M_1 - \beta_h}{x}$
due to the random choice of the mixture component $h$,
and (ii) the \emph{observation noise} $\epsilon \sim \sE$.
Although the noise depends on $x$,
it still has zero mean conditioned on $x$.
We will later show that we can
perform linear regression on the data $\{\xni,
\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$.
%under identifiability conditions.
But clearly, knowing $M_1$ is insufficient
for identifying all the parameters $\theta$,
as
$M_1$ only contains $d$ degrees of freedom whereas $\theta$ contains $O(kd)$.

@@

% Spectral
In recent years, there has a been a surge of interest in ``spectral'' methods
for learning latent-variable models.  One line of work has
focused on observable operator models \cite{hsu09spectral,song10kernel,parikh12spectral,cohen12pcfg,balle11transducer,balle12automata}
in which a reparametrization of the true parameters are recovered,
which suffices for prediction and density estimation.
Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
to recover the parameters \cite{anandkumar12moments,anandkumar12lda,hsu12identifiability,hsu13spherical}.
Our work extends this second line of work to models that
require regression to obtain the desired tensor.

>>>>

%%% first moment
To warm up, let us consider linear regression
on the response $y$ given $x$.
From the model definition, we have $y = \beta_h^\top x + \epsilon$.
The challenge is that the regression coefficients $\beta_h$ depend on the random $h$.
%$\epsilon \sim \normal{0}{\E[\epsilon^2]}$, and $h$ is a random quantity
%independent of $x$.
The first key step is to average over this randomness by defining
average regression coefficients
$M_1 \eqdef \sum_{h=1}^k \pi_h \beta_h$.
Now we can express $y$ as a linear function of $x$ with non-random coefficients $M_1$
plus a noise term $\eta_1(x)$:
\begin{align}
  y &= \innerp{M_1}{x} +
  \underbrace{(\innerp{\beta_h - M_1}{x} + \epsilon)}_{\eqdef \eta_1(x)}. \label{eqn:y1}
\end{align}
The noise $\eta_1(x)$ is the sum of two terms:
(i) the \emph{mixing noise} $\innerp{M_1 - \beta_h}{x}$
due to the random choice of the mixture component $h$,
and (ii) the \emph{observation noise} $\epsilon \sim \sE$.
Although the noise depends on $x$,
it still has zero mean conditioned on $x$.
We will later show that we can
perform linear regression on the data $\{\xni,
\yni\}_{i=1}^{n}$ to produce a consistent estimate of $M_1$.
%under identifiability conditions.
But clearly, knowing $M_1$ is insufficient
for identifying all the parameters $\theta$,
as
$M_1$ only contains $d$ degrees of freedom whereas $\theta$ contains $O(kd)$.

@@

% Spectral
In recent years, there has a been a surge of interest in ``spectral'' methods
for learning latent-variable models.  One line of work has
focused on observable operator models \cite{hsu09spectral,song10kernel,parikh12spectral,cohen12pcfg,balle11transducer,balle12automata}
in which a re-parametrization of the true parameters are recovered,
which suffices for prediction and density estimation.
Another line of work is based on the method of moments and uses eigendecomposition of a certain tensor
to recover the parameters \cite{anandkumar12moments,anandkumar12lda,hsu12identifiability,hsu13spherical}.
Our work extends this second line of work to models that
require regression to obtain the desired tensor.

### 2013-05-22 doc/icml2013/model.tex

The mixture of linear regressions model \citep{VieleTong2002} defines
a conditional distribution over a response $y \in \Re$
given covariates $x \in \Re^d$.
Let $k$ be the number of mixture components.
The generation of $y$ given $x$ involves three steps:
(i) draw a mixture component $h \in [k]$ according to mixture proportions
$\pi = (\pi_1, \dots, \pi_k)$;
(ii) draw observation noise $\epsilon$ from a known zero-mean noise distribution $\sE$,
and (ii) set $y$ deterministically based on $h$ and $\epsilon$.
More compactly: %to $\beta_h^\top x$ plus some observation noise $\epsilon$.
\begin{eqnarray}
  h &\sim& \mult(\pi), \\
  \epsilon &\sim& \sE, \\
  y &=& \beta_{h}^T x + \epsilon.
\end{eqnarray}
The parameters of the model are $\theta = (\pi, B)$,
where $\pi \in \Re^d$ are the mixture proportions and
$B = [\beta_1 \mid \dots \mid \beta_k] \in \Re^{d \times k}$
are the regression coefficients.
Note that the choice of mixture component $h$ and the observation noise $\epsilon$ are independent.
%We also assume the distribution of the observation noise is known and has bounded support.
The learning problem is stated as follows:
given $n$ i.i.d. samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
drawn from the model with some unknown parameters $\theta^*$,
return an estimate of the parameters $\hat\theta$.

% History of model
The mixture of linear regressions model has been applied
in the statistics literature for modeling music perception, where $x$ is the
actual tone and $y$ is the tone perceived by a musician \cite{VieleTong2002}.
% Reference mixture of experts.
% Acknowledge that mixture proportions % can't depend on $x$
The model is an instance of the hierarchical mixture of experts
\cite{jacobs91experts}, in which the mixture proportions are allowed to depend
on $x$, known as a gating function.
This dependence allow the experts to be localized in input space,
providing more flexibility, but we do not consider this dependence in our model.

>>>>

The mixture of linear regressions model \citep{VieleTong2002} defines
a conditional distribution over a response $y \in \Re$
given covariates $x \in \Re^d$.
Let $k$ be the number of mixture components.
The generation of $y$ given $x$ involves three steps:
(i) draw a mixture component $h \in [k]$ according to mixture proportions
$\pi = (\pi_1, \dots, \pi_k)$;
(ii) draw observation noise $\epsilon$ from a known zero-mean noise distribution $\sE$,
and (iii) set $y$ deterministically based on $h$ and $\epsilon$.
More compactly: %to $\beta_h^\top x$ plus some observation noise $\epsilon$.
\begin{eqnarray}
  h &\sim& \mult(\pi), \\
  \epsilon &\sim& \sE, \\
  y &=& \beta_{h}^T x + \epsilon.
\end{eqnarray}
The parameters of the model are $\theta = (\pi, B)$,
where $\pi \in \Re^d$ are the mixture proportions and
$B = [\beta_1 \mid \dots \mid \beta_k] \in \Re^{d \times k}$
are the regression coefficients.
Note that the choice of mixture component $h$ and the observation noise $\epsilon$ are independent.
%We also assume the distribution of the observation noise is known and has bounded support.
The learning problem is stated as follows:
given $n$ i.i.d.\ samples $(\xn{1}, \yn{1}), \dots, (\xn{n}, \yn{n})$
drawn from the model with some unknown parameters $\theta^*$,
return an estimate of the parameters $\hat\theta$.

% History of model
The mixture of linear regressions model has been applied
in the statistics literature for modelling music perception, where $x$ is the
actual tone and $y$ is the tone perceived by a musician \cite{VieleTong2002}.
% Reference mixture of experts.
% Acknowledge that mixture proportions % can't depend on $x$
The model is an instance of the hierarchical mixture of experts
\cite{jacobs91experts}, in which the mixture proportions are allowed to depend
on $x$, known as a gating function.
This dependence allow the experts to be localized in input space,
providing more flexibility, but we do not consider this dependence in our model.
